---
title: "Lab7"
author: "Dion, Julie, Mia, Vincent"
date: "9/20/2019"
output: 
    prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
data <- read.csv("Mall_customers.csv")
names(data)[4:5] <- c("Annual.Income", "Spending.Score")
```


```{r}
```

- function that computes the entropy --> conditional entropy, IG

# Exercise 1: Compute the information gain in the outcome due to knowing whether someone is older than 36 or not.


# Exercise 2: Compute the information gain in the outcome due to knowing whether someone’s annual income is above $50,000 or not.


# Exercise 3: Are the information gains in Exercises 2 and 3 like you would have expected? Why or why not? Could EDA have avoided any surprises here?


# Exercise 4: Perform k-means clustering on age, annual income, and the spending score. Explain your choice of k, visualize your results, and interpret the clusters. Make sure your analysis does not depend on the unit of measurements of the variables.

We start by subsetting the data and standardize each variable. Then we plot WSS against the number of clusters to determine a optimal value for k based on the plot. 
```{r}
kmdata <-  as.data.frame(data[c("Age", "Annual.Income", "Spending.Score")])
kmdata$Age <- kmdata$Age/(sd(kmdata$Age))
kmdata$Annual.Income <- kmdata$Annual.Income/(sd(kmdata$Annual.Income))
kmdata$Spending.Score <- kmdata$Spending.Score/(sd(kmdata$Spending.Score))

wss <- numeric(15)
for (k in 1:15) 
  wss[k] <- sum(kmeans(kmdata,centers=k, nstart=25)$withinss)
plot(1:15, wss, type="b", xlab="Number of Clusters", ylab="WSS")
```

k=4 seems to be the elbow of the WSS plot. So we choose k=4 to run the kmean analysis.  We run the kmean analysis below. 

```{r}
km <- kmeans(kmdata,4, nstart=25)
km
```

```{r}
library(ggplot2)
df = as.data.frame(kmdata)
df$cluster = factor(km$cluster)
centers=as.data.frame(km$centers)

g1 = ggplot(data=df, aes(x=Age, y=Annual.Income, color=cluster )) +
  geom_point() + theme(legend.position="right") +
  geom_point(data=centers,
             aes(x=Age,y=Annual.Income, color=as.factor(c(1,2,3,4))),
             size=10, alpha=.3, show.legend =FALSE)

g2 =ggplot(data=df, aes(x=Age, y=Spending.Score, color=cluster )) +
  geom_point() +
  geom_point(data=centers,
             aes(x=Age,y=Spending.Score, color=as.factor(c(1,2,3,4))),
             size=10, alpha=.3, show.legend=FALSE)

g3 = ggplot(data=df, aes(x=Annual.Income, y=Spending.Score, color=cluster )) +
  geom_point() +
  geom_point(data=centers,
             aes(x=Annual.Income,y=Spending.Score, color=as.factor(c(1,2,3,4))),
             size=10, alpha=.3, show.legend=FALSE)
g1
g2
g3
```


# Exercise 5: How does your k-means analysis compare to Section 3 (Segmentation using Age , Annual Income and Spending Score) from https://www.kaggle.com/kushal1996/customer-segmentation-k-means-analysis?

It show pretty different results, which mainly arise from a different choice of k. In this article, a k=6 is chosen. 


